Data Management for LSST Special Programs

### 2023-10-18 sec_proc.tex
update boundaries

\section{Processing for Special Programs Data Products}\label{sec:proc}

Rubin Observatory and the LSST Data Management System (DMS) are required 
to process Special Programs data to produce unique and separate data 
products whenever the original or reconfigured versions of the LSST 
Science Pipelines can be used to do so (Section~\ref{sec:req}).

The LSST Science Pipelines and the DMS are designed to process standard 
(or alternative standard) visits.

It depends on the science goals related to the Special Program, but some 
Special Programs' data can and should be processed and served along with 
(or as a part of) the WFD data products (Section~\ref{ssec:proc_wfd}).
Others will require RDM to generate unique and separate data products 
(Section~\ref{ssec:proc_rdm}).

Cases where the Special Program obtains non-standard visits that cannot 
be processed by the DMS (Section~\ref{ssec:proc_bounds}), and/or where 
development of new algorithms or the allocation of significant additional 
computational resources are needed, will require users tp generate unique
and separate data products (Section~\ref{ssec:proc_user}).

@@

Most of the Special Programs that are currently under consideration will 
use standard or alternative standard visits.

However some Special Programs are likely to use non-standard visits with 
shorter exposures (5 sec), and some are likely to acquire images with a 
significantly brighter sky background (e.g., twilight images) than 
standard (or altnerative) WFD main survey visits.

The cadence and patterns of Special Programs might also differ from the 
WFD, such as long series of exposures obtained of the same field 
(e.g., DDFs), or a strategy optimized to find very fast-moving objects 
(e.g., NEOs).

None of the currently proposed Special Programs listed in the Phase 2 
SCOC recommendations \citedsp{PSTN-055} 
seem likely to approach (or violate) any harware boundaries 
(Section~\ref{ssec:proc_bounds_hardware}).

However, some Special Programs observations seem likely to generate 
images that cannot be processed with original or reconfigured versions of 
the LSST Science Pipelines (Section~\ref{ssec:proc_bounds_processing}), 
and will require user processing.

@@

The minimum exposure time is 1 second (stretch goal: 0.1 seconds), and 
there is a potential hardware boundary that limits the readout rate to 1 
every 15 seconds.
This would affect the image acquisition rate and increase the overheads 
on short exposures.

Special Programs which required the \emph{exact same} field pointing and 
rotation for \emph{every exposure} (to sub-arcsecond levels) might run 
into hardware boundaries on pointing and tracking.
None of the currently-proposed programs seem likely to be affected by 
this potential boundary.

Hardware imposes no other boundaries on how data can be obtained, but 
Special Programs that request a high number of filter changes and/or long 
slews could be inefficient due to large overheads.

@@

Appendix~\ref{sec:procbounds} describes the boundaries on what types of visits 
can be processed and calibrated by the DMS.

Very short ($<$2 sec) exposures could be difficult to reduce due to an 
incompletely-formed PSF, and very short or very long ($>$150 sec) 
exposures could be difficult to calibrate due to having too few (or too 
few unsaturated) stars.
As mentioned above the DMS is required to be able to process exposure 
times as low as 1 second, and it is known that such short exposures might 
have degraded image quality.

It is currently unclear whether images with very bright sky backgrounds 
(twilight images) can be processed with the LSST Science Pipelines, or 
whether user generated pipelines will be needed.

The full reduction and calibration of images obtained with non-sidereal 
tracking, in which the stars are streaked, is currently beyond the scope 
of the LSST Science Pipelines, and will require a user generated pipeline.

>>>>

\section{Processing Data from Special Programs}\label{sec:proc}

% Rubin Observatory and the LSST Data Management System (DMS) are required 
% to process Special Programs data to produce unique and separate data 
% products whenever the original or reconfigured versions of the LSST 
% Science Pipelines can be used to do so (Section~\ref{sec:req}).

% The LSST Science Pipelines and the DMS are designed to process standard 
% (or alternative standard) visits.

% It depends on the science goals related to the Special Program, but some 
% Special Programs' data can and should be processed and served along with 
% (or as a part of) the WFD data products (Section~\ref{ssec:proc_wfd}).
% Others will require RDM to generate unique and separate data products 
% (Section~\ref{ssec:proc_rdm}).

% Cases where the Special Program obtains non-standard visits that cannot 
% be processed by the DMS (Section~\ref{ssec:proc_bounds}), and/or where 
% development of new algorithms or the allocation of significant additional 
% computational resources are needed, will require users to generate unique
% and separate data products (Section~\ref{ssec:proc_user}).

@@

Special Programs that do not use standard or alternative standard visits
might be limited by hardware or processing boundaries.

@@

In general, the currently-proposed Special Programs in \citeds{PSTN-055}
are not anticipated to be limited by hardware boundaries.

A few potential challenges to Special Programs posed by hardware
boundaries are discussed below.

\begin{itemize}

\item Special Programs that use short exposures would be limited to the
minimum exposure time is 1 second (stretch goal: 0.1 seconds).
There is a potential hardware boundary that limits the readout rate to 1 
every 15 seconds, which would affect the image acquisition rate and 
increase the overheads on short exposures.

\item Special Programs which require the \emph{exact same} field pointing and 
rotation for \emph{every exposure} (to sub-arcsecond levels) might run 
into hardware boundaries on pointing and tracking.

\item Special Programs that request a high number of filter changes and/or long 
slews could be inefficient due to large overheads, but there are no
boundaries on how data can be obtained.

\item Special Programs that obtain twilight images will be subject to safe
limits on sky background flux, as with any astronomical camera.

\end{itemize}

@@

Appendix~\ref{sec:procbounds} describes the boundaries on what types of visits 
can be processed and calibrated by the Data Management System and the LSST
Science Pipelines.

Most of the currently-proposed Special Programs in \citeds{PSTN-055}
are not anticipated to be limited by hardware boundaries.
However, those which use non-standard visits, especially those with
short exposure time or those obtained during twilight, might
be affected by processing boundaries.

The most likely challenges to Special Programs posed by processing
boundaries are discussed below.

\begin{itemize}

\item Special Programs which use very short ($<$2 sec) exposures 
could be difficult to reduce due to an incompletely-formed PSF 
(Section~\ref{ssec:procbounds_expt}).
The Data Management System is required to be able to process exposure 
times as low as 1 second, but it is known that such short exposures might 
have degraded image quality.

\item Special Programs that use very short or very long ($>$150 sec) 
exposures could be difficult to calibrate due to having too few 
(or too few unsaturated) stars.

\item For Special Programs which obtain images with very bright sky backgrounds
(twilight images), it is currently unclear whether they can be processed
with the LSST Science Pipelines; User-Generated Processing might be needed.

\item The full reduction and calibration of data from any Special Programs that 
use non-sidereal tracking, which produce images with star streaks, is
currently beyond the scope of the LSST Science Pipelines; 
User-Generated Processing would be needed.

\end{itemize}
